Transmission Effects On MPG In Cars
====================================

```{r data,echo=FALSE}
data(mtcars)
mtcarst <- within(mtcars, {
am <- factor(am, labels = c("Automatic", "Manual"))
vs <- factor(vs, labels = c("Vengine", "Iengine"))
cyl <- factor(cyl)
})
```

## 1. Executive Summary
This paper is based at 1974 Motor Trend Data that contains fuel consumption related with 10 aspects of automobile design and performance for 32 cars models 1973-74. The purpose is to estimate the effect of the type of transmission (automatic vs manual) on the miles per gallon (MPG). The results suggest that manual transmission have a higher MPG. However the weight and acceleration are attibutes that have a siginficant influence in the results.

## 2. Exploratory Data Analysis
Based on common sense we can assume that the mpg of a car depends on its weight and its power, also the cars with more cylinders have a great acceleration and more horse power. The following plot shows these theories.

```{r, echo=FALSE,fig.height=3,fig.width=10}
p<-c(16,17)
c<-c(4,6,8)
cl <- c("4 cyl.","6 cyl.","8 cyl.")
tl <- c("Automatic","Manual")

par(mfcol=c(1,2), mar=c(3.5,3.5,0,0.5), oma=c(0.5,5.5,0.5,0.5),cex=1,mgp=c(1.7,0.5,0))
with(mtcars,plot(mpg~wt,col=cyl,pch=am+p[1],xlab="Weight (lb/1000)",ylab="Miles/(US) gallon",main=""))
legend("topright",pch=p,legend=tl,bty="n")
legend("bottomleft",text.col=c,legend=cl,bty="n")

with(mtcars,plot(hp~qsec,col=cyl,pch=am+p[1],xlab="Acceleration (qsec)",ylab="Gross Horsepower",main=""))
legend("topright",pch=p,legend=tl,bty="n")
legend("bottomleft",text.col=c,legend=cl,bty="n")
```

The left plot offer following insights:
* The cars with automatic transmission are heavy.
* The cars with manual transmission are lightweight.
* A positive relationship between number of cylinders and weight.
* A negative relationship between MPG and weight.

The right plot offer following insights:
* A strong relationship between acceleration and horse power.
* Cars with more cylinders has great acceleration in lest time.
* Cars with more cylinders provides major horsepower.
* Cars with more horsepower have automatic transmission.

Also see the correlation matrix in the Appendix.

## 3. Models
The exploratory analysis suggested to use the wt variable with some power variables (cyl, hp) is a good start. An interesting insight is the acceleration that is strongly influenced by 8 cylinders cars therefore we will evaluate it. The models to analyze are:

```{r models, echo=FALSE}
model1 <- lm(mpg ~ am : wt + am : cyl, mtcarst) 
model2 <- lm(mpg ~ am : wt + am : hp, mtcarst)
model3 <- lm(mpg ~ am : wt + am : qsec, mtcarst)
formula(model1)
formula(model2)
formula(model3)
```

See the Appendix for results of model 1 and model 2. The stimation results reveal that model 3 is the best fit according following:

```{r model3}
summary(model3)
```

### 3.1 Model Validation
The model has a 88% Adjusted R-Squared, and contains three variables (transmission type, weight, and acceleration) along with the transmission interaction. The residuals are normal with 0 mean and constant variance. The "Residual vs Fitted" and "Scale Location" charts show that there is no trend to the residuals. The Q-Q plot shows that the errors are aproximately distributed. The p-values for am are considered statistically significant.

### 3.2 Model Interpretation
**For constant quarter-mile-time:** For each unit of increase in quarter-mile-time (1 unit = 1 sec), there is a 0.834 mpg increase for automatic cars while there is a 1.446 mpg increase for manual cars. This means manual cars have more efficient acceleration than automatic but in both cases require more fuel. 

**For constant weight (lb/1000):** For each unit of increase in weight (1 unit = 1000 lb) there is 3.176 decrease for automatic cars while there is a 6.099 mpg decrease for manual cars. This means heavier cars in general require more fuel altought automatic cars consumes less fuel than manual.



## Appendix
### Data
The data frame consists of 32 observations on 11 variables.
- mpg: Miles/(US) gallon
- cyl: Number of cylinders
- disp: Displacement (cu.in.)
- hp: Gross horsepower
- drat: Rear axle ratio
- wt: Weight (lb/1000)
- qsec: 1/4 mile time
- vs: V/S
- am: Transmission (0 = automatic, 1 = manual)
- gear: Number of forward gears

```{r sample,echo=FALSE}
head(mtcarst)
```

### MPG against all variables
```{r scatter}
require(graphics)
panel.cor <- function(x, y, digits=2, prefix="", cex.cor, ...) {
usr <- par("usr"); on.exit(par(usr))
par(usr = c(0, 1, 0, 1))
r <- cor(x, y)
txt <- format(c(r, 0.123456789), digits=digits)[1]
txt <- paste(prefix, txt, sep="")
if(missing(cex.cor)) cex.cor <- 0.8/strwidth(txt)
text(0.5, 0.5, txt, cex = cex.cor * abs(r))
}
pairs(mpg~am:., data=mtcarst, lower.panel=panel.smooth, upper.panel=panel.cor, pch=20, main="Scatterplot - Correlation Matrix")

```

### Models
**Model 1**
```{r model1}
summary(model1)
```

**Model 2**
```{r model2}
summary(model2)
```

**Selected Model Plot**
```{r modelval}
par(mfrow=c(2,2),cex.main=0.1,cex=0.6,mgp=c(1.7,0.5,0),mar=c(4,3.5,2,0.5),oma=c(0.5,10,0.5,0.5))
plot(model3)
```